Parameter precision in economic growth models inspired by Ecological Economics

### 2015-10-05 Articles/Article_EcolEcon_2015/Paper0/Paper0-Resampling.Rnw
Edits to models and resampling sections.

The Constant Elasticity of Substitution~(CES) model 
is given as
%
\begin{equation} \label{eq:CESgeneric}
  y = \gamma \: A \: \left\{\delta \left[\delta_1 x_1^{-\rho_1} 
      + (1-\delta_1)x_2^{-\rho_1} \right]^{\rho/\rho_1} 
      + (1-\delta) x_3^{-\rho} \right\}^{-1/\rho}; 
      A \equiv \mathrm{e}^{\lambda (t-t_0)},
\end{equation}
%
where
$x_1$, $x_2$, and $x_3$ are factors of production
and permutations of capital~($k$), labor~($l$), and energy~($e$).

For the CES function, 
output elasticities~($\alpha_1$, $\alpha_2$, and $\alpha_3$) are not constant;
rather, they vary with factors of production ($x_1$, $x_2$, and $x_3$) over time.
%
\begin{equation} \label{eq:alpha_x_1}
  \alpha_{x_1} \equiv
  \frac{ \frac{\partial y}{\partial x_1} }
          { \frac{y}{x_1} } 
  = 
  \frac{ \delta  \delta _1 x_1^{-\rho _1} \left[\delta _1 x_1^{-\rho _1} + \left(1 - \delta_1 \right) x_2^{-\rho _1}\right]
                             ^{\frac{\rho}{\rho _1}-1} }
  { \delta \left[\delta _1 x_1^{-\rho _1} + \left(1 - \delta _1 \right) x_2^{-\rho _1} \right]
                             ^{ \frac{\rho }{\rho_1} } 
                + (1 - \delta) x_3^{-\rho }}
\end{equation}
%
\begin{equation} \label{eq:alpha_x_2}
  \alpha_{x_2} \equiv \frac{ \frac{\partial y}{\partial x_2} }{\frac{y}{x_2}} =
  \frac{ \delta  \left(1-\delta _1 \right) x_2^{-\rho _1} \left[\delta _1 x_1^{-\rho _1}+\left(1 - \delta_1\right) x_2^{-\rho_1}\right]
	 				^{\frac{\rho }{\rho _1}-1} }
		{\delta  \left[\delta _1 x_1^{-\rho_1} + \left(1 - \delta_1\right) x_2
                                        ^{-\rho_1}\right]^{\frac{\rho }{\rho _1}}+ (1 - \delta) {x_3}^{-\rho }}
\end{equation}
%
\begin{equation} \label{eq:alpha_x_3}
  \alpha_{x_3} \equiv \frac{ \frac{\partial y}{\partial x_3} }{ \frac{y}{x_3} } =  
      \frac{1 - \delta}{\delta  x_3^{\rho } \left[\delta _1 x_1^{-\rho_1} + \left(1 - \delta_1 \right) x_2^{-\rho _1}\right]
				^{\frac{\rho }{\rho_1}} + 1 - \delta}
\end{equation}
%
It can be verified that
%
\begin{equation} \label{eq:sum_betas}
  \alpha_{x_1} + \alpha_{x_2} + \alpha_{x_3} = 1
\end{equation}
%
for all nests.

Elasticities of substitution~($\sigma$) between factors of production
are a function of $\rho_1$, $\rho$, and nesting. 
**** More here on the actual $\sigma$ equations. 
Gather them from the Henningsen and Henningsen paper. ****

@@

%------------------------------
\subsubsection{Technique} 
\label{sec:parameter_estimation_technique}
%------------------------------

To estimate values of parameters
($\theta$, $\lambda$, $\delta_1$, $\delta$, $\rho_1$, and $\rho$)
for the reference and CES models,
we use an ordinary least squares~(OLS) approach.
The objective of the OLS analysis is minimization of the sum of squared errors~(\sse):
%
\begin{equation} \label{eq:sse}
  \sse = \sum_{i} r_i^2 \; ,
\end{equation}
%
where
$r_i \equiv \ln (y_i / \hat{y}_i)$ is $i^{\mathrm{th}}$ residual 
(all models assume multiplicative errors),
and
$\hat{y}_i$ is the fitted value for economic output at time $t_i$.
In cases where a model adheres to the cost share principle,
output elasticities are fixed
in accordance with long-run historical cost shares
for capital and labor ($\hat{\alpha}_k = 0.3$ and $\hat{\alpha}_l = 0.7$).%
\authNoted{Randy put hats on 
$\alpha_k$ and $\alpha_l$, 
but I don't think they should be ``hatted.''
These are not estimated numbers. 
They are give \emph{a-priori} from national accounts data. ---MKH;

This is an edge case.  From data = estimated; just estimated in a different way,
and external to the data we are analyzing.  But that doesn't make those values
correct. ---rjp

OK. I'm willing to live with the hats. ---MKH}
We estimate parameters using a highly-customized version
of the \texttt{R} package \texttt{micEconCES} \citep{Henningsen:2011td}.
See the Supplemental Information
(\ref{sec:appendix_CES_fitting}) for additional details.
\authNoted{Is there a reason to use \mse{} rather than \sse?  Typically,
the denominator for \mse{} is not $N$ but the degrees of freedom
used to estimate the variance parameter of the model (typically $N - p$ for
a model with $p$ parameters).  Perhaps we should refer to this as
an ``unadjusted'' \mse{} if we stick with it.
I changed everything to sse. ---MKH}

\authNoted{What sort of ``expect'' is this? Can we say something more precise? ---rjp
I changed to ``know,'' which is more precise. ---MKH}

In the reference model,
all economic growth is attributed to the Solow residual. %~($\lambda$).
Indeed, the reference model will
have a larger estimated Solow 
residual~($\hat{\lambda}$) than
any CES model, because no factors of production are
included in the reference model to drive economic growth.
(If all of $y$, $k$, $l$, and $e$ are $\ge 1$,
the non-$A$ part of a CES model
will be $\ge 1$,
thereby driving $A$, and $\lambda$, to be smaller.)

\authNoted{I'm not sure this is true, and if it is, I think it is misleading, since
as $\rho \to 0$, CES $\to$ C-D, right?  Even if we disallow $\rho = 0$, this suggests
that the CES model can have \sse{} arbitrarily close to that for the C-D model, even if
the C-D model were best.  In practice, I'm guessing one will never see \sse{} for
CES exceeding that for C-D.  And in our CES function, we fit the limiting cases
explicitly, so it should be impossible.  Matt, am I forgetting something here? ---rjp
CD $\ne$ reference model.
The reference model is exponential-only. ---MKH}

On the other hand, it is not necessarily true that
CES models (Equations~\ref{eq:CESkl} and \ref{eq:CESkle}--\ref{eq:CESekl})
will exhibit lower \sse{} than the reference model
(Equation~\ref{eq:exponential-only-model}).
The CES production functions
have more parameters,
but there is no set 
of parameter 
estimates~($\hat\theta$, $\hat\lambda$, $\hat\delta_1$, 
$\hat\delta$, $\hat\rho_1$, and $\hat\rho$)
that mathematically eliminates 
the factors of production~($k$, $l$, and $e$)
from the CES models, 
thereby reproducing the reference model.%
\authNote{I added the clause ``but there is no combination of parameters
that can eliminate the factors of production
from the models.'' Is Randy OK with this? ---MKH}
%but they incorporate the factors of production
%at constant returns to scale, and we constrain them so that 
%$\alpha_k + \alpha_l + \alpha_e = 1$.
%In contrast, the reference model has $\alpha_k = \alpha_l = \alpha_e = 0$.
%\authNoted{Have we defined $\alpha_i$ for the reference model? ---rjp
%I have now added a sentence stating that reference models are a degenerate 
%form of CES models with $\alpha_i = 0$. ---MKH
%But this still isn't true --rjp}%
If the factors of production
are poorly correlated to output~($y$),
\sse{} may be higher for a CES model than for the reference model.
\authNoted{rjp:  removed bit about CRS.  I don't think CRS is the reason.
That's fine. ---MKH}%

%++++++++++++++++++++++++++++++
\subsection{CES fitting algorithm}
\label{sec:appendix_CES_fitting}
%++++++++++++++++++++++++++++++

We use the \texttt{R}~\citep{R}
package \texttt{micEconCES}
\citep{Henningsen:2011td}
with significant customization
to estimate parameters in the CES production function.
Our CES parameter estimation algorithm starts with an eleven-value grid search
in $\rho_1$ and $\rho$
(9, 2, 1, 0.43, 0.25, 0.1, -0.1, -0.5, -0.75, -0.9, -0.99),
corresponding to $\sigma_1$ and $\sigma$ values of
0.1, 0.33, 0.5, 0.7, 0.8, 0.9, 1.11, 2, 4, 10, and 100,
respectively.
During the grid search, values of $\rho_1$ and $\rho$ are fixed,
and values of $\theta$, $\lambda$, $\delta_1$, and $\delta$ are estimated
by gradient search with the \texttt{PORT} and \texttt{L-BFGS-B} algorithms
in the \texttt{micEconCES} package~\citep{Henningsen:2011td}.
In all, 121 gradient searches in $\theta$, $\lambda$, $\delta_1$, and $\delta$
at grid points representing all combinations of $\rho_1$ and $\rho$ are
attempted.
During the grid search portion of our algorithm, starting values for the free
parameters
are $\lambda$ = 0.015/year, $\delta_1$ = 0.5, $\delta$ = 0.5,
and $\theta$ is set by the \texttt{cesEst} function
to a value such that the mean of the residuals is zero.

@@

Henningsen and Henningsen~\cite{Henningsen:2011td},
in their detailed analysis of Kemfert~\cite{kemfert1998estimated}, found that
%
\begin{quote}
\ldots the Levenberg-Marquardt and the \texttt{PORT} algorithms
are---at least in this study---most likely to find the
coefficients that give the best fit to the model,
where the \texttt{PORT} algorithm can be used to restrict the estimates
to the economically meaningful region.
\end{quote}
%
In our testing, we found that to be mostly true.
\texttt{PORT} nearly always provides lower
\sse{} than \texttt{L-BFGS-B}, despite the fact that
\texttt{L-BFGS-B} often reports convergence
and \texttt{PORT} does not for the same data
(i.e., for the same $y$, $k$, $l$, and $e$ time series).

@@

%------------------------------	439 words on 14-09-2015
\subsubsection{Precision} 
\label{sec:parameter_precision}
%------------------------------

Determining the precision of estimates for CES parameters
($\hat\theta$, $\hat\lambda$, $\hat\delta_1$, $\hat\delta$, $\hat\rho_1$, 
and $\hat\rho$, but also
$\hat\alpha_k$, $\hat\alpha_l$, $\hat\alpha_e$, 
$\hat\sigma_{kl}$, $\hat\sigma_{le}$, $\hat\sigma_{ek}$, 
$\hat\sigma_{(kl)e}$, $\hat\sigma_{(le)k}$, and $\hat\sigma_{(ek)l}$)
is important but challenging.
It is quite possible for substantial changes in a CES model parameter
to have a relatively modest effect on the objective function
that is determining the parameter estimates (in this case, \sse).
When this happens, the ``best estimate''
is not much better than many other good estimates, and parameter
estimates must be interpreted with caution.

The usual methods for quantifying the precision of parameter estimates
using standard errors, confidence intervals,
and $p$-values rely on an asymptotic theory that applies
on the interior of a parameter space and assumes independence of error terms.
% With $n \approx 50$\authNoted{We have 50 years, 1960--2009, not 30 years as Randy 
% originally thought. 
% Does this change the way we write this section? ---MKH
% Well, we're closer to infinity, but we still have edges --rjp} 
% (in this case, years)
% \authNoted{Do you mean 30 years?
% Should we rather say ``With 30 years of data''---MKH.
% Perhaps there is a better word than sample, but it isn't really primarily
% about time but about the number of rows of data.  That happens to equal the number
% of years of data in this case but isn't the real issue.
% For now, I'm punting and just using $N \approx 30$. ---rjp}%
% and with CES functions (Equations~\ref{eq:CESkl} and~\ref{eq:CESkle}--\ref{eq:CESekl})
% having as many as six parameters,
It can be difficult to provide an \emph{a priori} justification
for the use of such asymptotic results.
The authors of the \texttt{R} package \texttt{micEconCES}
acknowledge as much when they say that
``As [the computation of the variance-covariance
matrix] is only valid asymptotically,
we calculate the estimated variance of the residuals \dots
without correcting for degrees of freedom.'' \citep{Henningsen:2011td}

Further complicating matters is our interest in issues that arise on the
boundary of (the economically meaningful portion of) the parameter space.
In an investigation of whether energy is important, for example,
we may consider the null hypothesis that $\alpha_e = 0$.  
In the ($kl$)$e$ nesting, 
this null hypothesis is equivalent to $\delta = 1$.
While there is an asymptotic theory for this case as well,%
\authNote{%
Randy: can you provide a reference for the asymptotic theory? ---MKH}%
the distributions are, in general,
more complicated---mixtures of chi-squared variates---and the
precise mix may be difficult or impossible to determine analytically.
\authNote{Ideally, we should check that there is not an easy analytical solution
in our particular case.  We'll see if there is time. ---rjp}%

@@

%++++++++++++++++++++++++++++++
\subsection{Bootstrap resampling algorithm}
\label{sec:appendix_resampling}
%++++++++++++++++++++++++++++++

Bootstrapping is a statistical resampling technique for assigning
measures of accuracy and precision to sample estimates by measuring
properties of an estimator when sampling from a resampling
distribution.
Resampling distributions can be formed in a number of
ways in accordance with the type of data, experimental design,
and modeling assumptions involved.
In each case, many new resamples
are created, each of which is a randomized version of the original sample data to
which the desired analysis method can be applied.
By investigating, for example,
the variability of a parameter estimate from one resample to another,
one can learn about the precision of the estimation method.
**** Add some references here. ---rjp ****
% \authNote{Add some references here. ---rjp}

In the context of linear models (regression),
resamples are generally created
by residual resampling.
In our case, we formed resamples by adding to the fitted
response~($\hat{y}$) the product of a residual from the original CES function fit
(see Figure~\ref{fig:r_kle}) and
random sign ($-1$ or $1$, each with probability $0.5$).
This residual resampling occurs on the
log-transformed data.
Intuitively, this method assumes that the residuals are indicative of
the variability (from many potential sources) inherent in the data such that
it would be unsurprising if the residual from
any particular year had been observed in a different year.
Thus, a resampled response $\tilde Y'$ can be computed as
**** Randy edit the following equations to reflect the fact that
we're fitting in log space? ---MKH ****
%
\begin{equation}
  \tilde Y_i' = \hat y_i \pm r_j \; ,
\end{equation}
%
where
%
\begin{equation}
  r_i = y_i - \hat y_i \; .
\end{equation}
%
Both the sign~($\pm$) and the index
($j$,~typically different from $i$)
of the residual~($r_j$)
are chosen at random (with replacement).
We repeated the resampling process 1000 times for each
combination of growth model and country,
both with and without energy as appropriate.

>>>>

The Constant Elasticity of Substitution~(CES) model 
is given as
%
\begin{equation} \label{eq:CESgeneric}
  y = \theta \: A \: \left\{\delta B_1^{\rho/\rho_1} 
      + (1-\delta) x_3^{-\rho} \right\}^{-1/\rho}; 
      A \equiv \mathrm{e}^{\lambda t},
\end{equation}
%
where
$B_1 \equiv \delta_1 x_1^{-\rho_1} + (1-\delta_1) x_2^{-\rho_1}$, and
$x_1$, $x_2$, and $x_3$ are factors of production
and permutations of capital~($k$), labor~($l$), and energy~($e$).

For the CES function, 
output elasticities~($\alpha_1$, $\alpha_2$, and $\alpha_3$) are not constant;
rather, they vary with factors of production ($x_1$, $x_2$, and $x_3$) over time.
%
\begin{equation} \label{eq:alpha_1}
  \alpha_1 \equiv
  \frac{ \frac{\partial y}{\partial x_1} }
          { \frac{y}{x_1} } 
  = 
  \frac{ \delta  \delta _1 x_1^{-\rho _1} B_1^{ ( \frac{\rho}{\rho_1}-1 ) }  }
  { \delta B_1^{ ( \frac{\rho }{\rho_1} ) } + (1 - \delta) x_3^{-\rho }}
\end{equation}
%

\begin{equation} \label{eq:alpha_2}
  \alpha_2 \equiv \frac{ \frac{\partial y}{\partial x_2} }{\frac{y}{x_2}} =
  \frac{ \delta  \left(1-\delta _1 \right) x_2^{-\rho _1} B_1^{ ( \frac{\rho }{\rho _1}-1 ) } }
  { \delta B_1^{ ( \frac{\rho }{\rho_1} ) } + (1 - \delta) x_3^{-\rho }}
\end{equation}
%
\begin{equation} \label{eq:alpha_3}
  \alpha_3 \equiv \frac{ \frac{\partial y}{\partial x_3} }{ \frac{y}{x_3} } =  
      \frac{1 - \delta}{\delta  x_3^{\rho } B_1^{(\frac{\rho }{\rho_1} ) } + 1 - \delta}
\end{equation}
%
It can be verified that
%
\begin{equation} \label{eq:sum_betas}
  \alpha_1 + \alpha_2 + \alpha_3 = 1
\end{equation}
%
for all nests.

Similar to the output elasticities, 
the elasticity of substitution between factors of production $x_1$ and $x_2$~($\sigma_{1,2}$) 
is time-varying and given by the Allen-Uzawa elasticity of substitution:
%
\begin{equation} \label{eq:sigma_1_2}
  \sigma_{1,2} = \frac{ (1 - \rho_1)^{-1} - (1-\rho)^{-1}}{\delta \left( \frac{y}{B_1^{-\frac{1}{\rho_1}}} \right) ^ {1 + \rho}}
                      + (1 + \rho)^{-1} \; .
\end{equation}
%

**** Add Hicks-McFadden elasticity of substitution here. ****

@@

%------------------------------
\subsubsection{Cobb-Douglas model} 
\label{sec:param_estimation_CD}
%------------------------------

To estimate the parameters $\theta$, $\lambda$, $\alpha_1$, $\alpha_2$, and $\alpha_3$ 
in the Cobb-Douglas model, 
we reparameterized Equation~\ref{eq:CD} as
%
\begin{equation} \label{eq:CDe_AB_reparam}
  y = \theta \, \mathrm{e}^{\lambda t} \, x_1^{\alpha_1} \, x_2^{\alpha_2} \, x_3^{1 - \alpha_1 - \alpha_2}
\end{equation}
%
ensuring that $\alpha_1 + \alpha_2 + \alpha_3 = 1$,
thereby providing constant returns to scale.
The \texttt{R} \citep{R} function \texttt{lm} was used to estimate 
values of $\lambda$, $\theta$, $\alpha_1$, and $\alpha_2$ in log-transform space.
If the fitted value for $\alpha_1$ or $\alpha_2$ fell outside the interval $[0,1]$,
we fit along all boundaries 
($\alpha_1 = 1$,
$\alpha_2 = 1$,
$\alpha_3 = 1$,
$\alpha_1 = 0$,
$\alpha_2 = 0$, and
$\alpha_3 = 0$)
and chose the boundary fit with
minimum sum of squared errors~(\sse) as the winning model:
%
\begin{equation} \label{eq:sse}
  \sse = \sum_{i} r_i^2 \; ,
\end{equation}
%
where the residual is given as
%
\begin{equation} \label{eq:residual}
  r_i \equiv \ln (y_i / \hat{y}_i)
\end{equation}
%
is the $i^{\mathrm{th}}$ residual (all models assume multiplicative errors),
and
$\hat{y}_i$ is the fitted value for economic output at time $t_i$.
$\alpha_3$ was recovered with $\alpha_3 = 1 - \alpha_1 - \alpha_2$.

In cases where a model adheres to the cost share principle,
output elasticities are fixed
in accordance with long-run historical cost shares
for capital and labor ($\hat{\alpha}_1 = 0.3$ and $\hat{\alpha}_2 = 0.7$).
Under these conditions, $\hat{\alpha}_3$ is assumed to be $0$.

%------------------------------
\subsubsection{CES model} 
\label{sec:param_estimation_CES}
%------------------------------

To estimate values of parameters
($\theta$, $\lambda$, $\delta_1$, $\delta$, $\rho_1$, and $\rho$)
for the CES model,
we use an ordinary least squares~(OLS) approach.
The objective of the OLS analysis is minimization of the sum of squared errors~(Equation~\ref{eq:sse}).
We estimate parameters using the \texttt{R}~\citep{R} package \texttt{micEconCES}
\citep{Henningsen:2011td}
with significant customization
to estimate parameters in the CES production function.
Our CES parameter estimation algorithm starts with an eleven-value grid search
in $\rho_1$ and $\rho$
(9, 2, 1, 0.43, 0.25, 0.1, -0.1, -0.5, -0.75, -0.9, -0.99),
corresponding to $\sigma_1$ and $\sigma$ values of
0.1, 0.33, 0.5, 0.7, 0.8, 0.9, 1.11, 2, 4, 10, and 100,
respectively.
During the grid search, values of $\rho_1$ and $\rho$ are fixed,
and values of $\theta$, $\lambda$, $\delta_1$, and $\delta$ are estimated
by gradient search with the \texttt{PORT} and \texttt{L-BFGS-B} algorithms
in the \texttt{micEconCES} package~\citep{Henningsen:2011td}.
In all, 121 gradient searches in $\theta$, $\lambda$, $\delta_1$, and $\delta$
at grid points representing all combinations of $\rho_1$ and $\rho$ are
attempted.
During the grid search portion of our algorithm, starting values for the free
parameters
are $\lambda$ = 0.015/year, $\delta_1$ = 0.5, $\delta$ = 0.5,
and $\theta$ is set by the \texttt{cesEst} function
to a value such that the mean of the residuals is zero.

@@

\citet{Henningsen:2011td},
in their detailed analysis of Kemfert~\cite{kemfert1998estimated}, found that
%
\begin{quote}
\ldots the Levenberg-Marquardt and the \texttt{PORT} algorithms
are---at least in this study---most likely to find the
coefficients that give the best fit to the model,
where the \texttt{PORT} algorithm can be used to restrict the estimates
to the economically meaningful region.
\end{quote}
%
In our testing, we found that to be mostly true.
\texttt{PORT} nearly always provides lower
\sse{} than \texttt{L-BFGS-B}, despite the fact that
\texttt{L-BFGS-B} often reports convergence
and \texttt{PORT} does not for the same data
(i.e., for the same $y$, $k$, $l$, and $e$ time series).

@@

%++++++++++++++++++++++++++++++
\subsection{Parameter precision} 
\label{sec:parameter_precision}
%++++++++++++++++++++++++++++++

Determining the precision of estimates for 
Cobb-Douglas parameters
($\hat\theta$, $\hat\lambda$, $\hat\alpha_1$, $\hat\alpha_2$, and $\hat\alpha_3$)
and 
CES parameters
($\hat\theta$, $\hat\lambda$, $\hat\delta_1$, $\hat\delta$, $\hat\rho_1$, 
and $\hat\rho$, but also time series for
$\hat\alpha_1$, $\hat\alpha_2$, $\hat\alpha_3$, 
$\hat\sigma_{1,2}$, and $\hat\sigma_{(1,2)3}$)
is important but challenging.
It is quite possible for substantial changes in a model parameter
to have a relatively modest effect on the objective function
that is determining the parameter estimates (in this case, \sse).
When this happens, the ``best estimate''
is not much better than many other good estimates, and parameter
estimates must be interpreted with caution.

The usual methods for quantifying the precision of parameter estimates
using standard errors, confidence intervals,
and $p$-values rely on an asymptotic theory that applies
on the interior of a parameter space and assumes independence of error terms.
It can be difficult to provide an \emph{a priori} justification
for the use of such asymptotic results.
The authors of the \texttt{R} package \texttt{micEconCES}
acknowledge as much when they say that
``As [the computation of the variance-covariance
matrix] is only valid asymptotically,
we calculate the estimated variance of the residuals \dots
without correcting for degrees of freedom.'' \citep{Henningsen:2011td}

Further complicating matters is our interest in issues that arise on the
boundary of (the economically meaningful portion of) the parameter space.
In an investigation of whether energy is important, for example,
we may consider the null hypothesis that $\alpha_3 = 0$.  
In the ($kl$)$e$ nesting for the CES model, 
this null hypothesis is equivalent to $\delta = 1$.
While there is an asymptotic theory for this case as well,%
  \authNote{Randy: can you provide a reference for the asymptotic theory? ---MKH}
the distributions are, in general,
more complicated---mixtures of chi-squared variates---and the
precise mix may be difficult or impossible to determine analytically.
\authNote{Ideally, we should check that there is not an easy analytical solution
in our particular case.  We'll see if there is time. ---rjp}%

@@

Bootstrapping is a statistical resampling technique for assigning
measures of accuracy and precision to sample estimates by measuring
properties of an estimator when sampling from a resampling
distribution.
Resampling distributions can be formed in a number of
ways in accordance with the type of data, experimental design,
and modeling assumptions involved.
In each case, many new resamples
are created, each of which is a randomized version of the original sample data to
which the desired analysis method can be applied.
By investigating, for example,
the variability of a parameter estimate from one resample to another,
one can learn about the precision of the estimation method.
\authNote{Add some references here. ---rjp}

In the context of linear models (regression),
resamples are generally created
by residual resampling.
In our case, we formed resamples by adding to the fitted
response~($\hat{y}$) the product of a residual from the original CES function fit
(see Figure~\ref{fig:r_kle}) and
random sign ($-1$ or $1$, each with probability $0.5$).
This residual resampling occurs on the
log-transformed data.
Intuitively, this method assumes that the residuals are indicative of
the variability (from many potential sources) inherent in the data such that
it would be unsurprising if the residual from
any particular year had been observed in a different year.
Thus, a resampled response $\tilde Y'$ can be computed as%
\authNote{Randy: edit the following equations to reflect the fact that
we're fitting in log space? ---MKH}

\begin{equation} \label{eq:resampled_y}
  \tilde Y_i' = \hat y_i \pm r_j \; ,
\end{equation}
%
where
%
\authNote{Do we need this equation for $r_i$ here? We already have Equation~\ref{eq:residual}.}%
%
\begin{equation} \label{eq:residual_2}
  r_i = y_i - \hat y_i \; .
\end{equation}
%
Both the sign~($\pm$) and the index
($j$,~typically different from $i$)
of the residual~($r_j$)
are chosen at random (with replacement).
We repeated the resampling process 1000 times for each
combination of growth model and country,
both with and without energy as appropriate.
